PLANETESIMAL ACCRETION AT SHORT ORBITAL PERIODS

### 2023-04-21 refereeResponse.txt

2. Overview of planetesimal accretion
2.1 Oligarchic and Runaway Growth
144 onwards: indices in math upright
	Fixed
152: Define rms velocity
	TODO: not sure what the referee wants? Its defined in equation 1?
161: What does "orderly" fashion mean exactly?
	This means the mass ratios between bodies tend toward unity. This has been clarified in the text.
174: Add reference.
	Added reference to Lissauer 1987. Also mentioned that pebble accretion is a viable alternative.
198: Ln Lambda approx 10 missing could use a reference.
	Typo in the text, Ln Lambda ~ 3 is typically used. Added references to Ida 1990, Stewart and Ida 1993). Ln Lambda = 3 was used to construct figure 1, so no changes to the figure needed.
204: I suggest more formally introducing the concept of Hill sphere and hill factor.
	TODO: Where to put this?
209: This sentence seems not necessary.
	Agreed, removed
213: add "dynamically" heating
	Added
222: add "a" crucial assumption
	Fixed
226: Maybe clarify that growth timescale equals collision time scale.
	We have replaced 'growth' with 'collision' here, and done the same elsewhere in the paper for consistency
238: The eccentricity dispersion e_h is never defined anywhere. It is unclear to me what scaling e_h with the Hill-factor means physically.
	e_h can be thought of as the radial excursion of the body's orbit in units of its Hill radius. For e_h > 1, the random velocity of a body dominates over the velocity difference due to the shearing of the disk (over a separation of 1 r_h) and encounters between bodies can be treated with a two-body formalism. We have decided to remove any mention of e_h here, and instead save this definition for section 4 (TODO: define in section 4), where it is next mentioned.
Figure .1:
Since this is an analytical plot, I think one could color the entire parameter space according to the ratio t_relax/t_coll. Currently, it may suggest that the plotted values for e_h are chosen for a particular reason.
	The y axis now represents e_h, and the space on the plot is colored by t_relax/t_coll.
Also, I think it would be nice to have a reference for what values of e_h are standard for certain disk conditions, i.e. how outrageous would it be to have a e_h = 16? Are all disks like that? Or do we expect e_h = 1 as the standard?
	e_h of order ~10 is pretty standard. Following Kokubo 2002, we choose e_h so that the viscous stirring and gas drag timescales are equal. The equilibirum value of e_h then depends on the local solid and gas density, and will vary a bit with orbital period. In all of the simulations presented in this paper, the equilibrium value for e_h ranges between ~5 and ~50. We have overplotted some curves in figure 1 to show the equilibrium value of e_h as a function of orbital period for various values of the dust-gas ratio (assuming a MMSN solid surface density profile). TODO: Center white on colormap at 1
Also, I recommend adjusting the x-limits of the plot to correspond to the limits of the orbital period.
	Done

@@

702: Are they that similar? Maybe this statement can be more quantitative? To me it looks like the values fall below 2sqrt(3) at 40 days for fdHiShallow and at 70 days for fdLo. The interpretation that the boundary is only driven by the local value for alpha seems somewhat speculative. I'd also be interested how randomness affects this, i.e. if the simulation is initialized with a different random seed, by how much does the final distribution change?
	We were saving these for the follow up paper, but we do have multiple versions of fdHi, fdSteep and fdShallow! These have now been added to figure 8. The added data does make it clear that the boundary is quite fuzzy. I've also plotted the mean and standard deviation of the \tilde{b} in a number of orbital period bins. I think this makes the accretion boundary a little more apparent. With the exception of maybe the fdShallow simulation, the locations in the disk where \tilde{b} drops below 2\sqrt{3} are broadly consistent, which supports our claim that the boundary location is driven by the local value of alpha.
	TODO: Update text to recflect changes to figure.

5.5. Assembly History of Embryos
719: add references.
	Done
721: Does this distinction have analogies in galaxy formation? Also isn't this overly strict? I would expect that the initial population has quickly collided at least once, so most of the collisions will be major mergers under this definition. This seems to be supported by Fig. 9 where the smooth accretion fraction is always <10%.
	In galaxy formation simulations, 'smooth' accretion is considered to be anything at or below the resolution limit. Your point about the smooth accretion fraction always being small due to this overly strict cutoff is a good one. We have reconstructed figure 9 using a mass cutoff of 100 m_0 for major/minor collisions.
724: Why are only the 100 largest bodies chosen? It seems more natural to me to choose a mass cut-off. Also, given that in Fig9 is plotting mass on a log-scale, I suggest just plotting the entire population including those planetesimals that remained relatively small and pristine. Also, why is the fdHi simulation chosen?
	The 100 body cutoff was chosen because the analysis script we were using the construct the collision history was rather slow when asked to include all ~10,000 particles in the snapshot. This has been optimized and all bodies larger than 100 m_0 are now included in figure 9. fdHi is chosen to be the 'fiducial' simulation for most of this paper. The initial conditions for fdHi require the least assumptions (MMSN profile with enough extra mass included to construct ~1 M_earth planets).
726: I think it could be explicitly stated that smooth accretion = minor mergers.
	TODO: How to address this?
730 & onwards: Can the authors quantify this trend? E.g., by making three linear fits (with error bars) for > 60d, < 60d and the full population. I'm in particular unconvinced that the trend for P > 60 days truly is flat, given the small sample size for larger bodies.
	As with figure 8, we have now included data from the other fdHi runs (using a different random seed). The P > 60d trend is now much more clearly flat.
Figure 9:
Given that I'm skeptical about the 60 day boundary being clear and universal, I think this plot would be stronger if the bodies were colored by orbital period (in addition to the marker style indicating this boundary), and the plot would include a corresponding colorbar.
	The mass cut has been increased (100 m_0), points are now colored by orbital period and the 60d boundary is indicated by a different marker style. 
The authors state "there is a decreasing correlation between smooth accretion fraction and mass, which suggests that most growth occurs between equal mass bodies". As far as I understand, this doesn't have to be the case. I would argue that more massive bodies require more time to build up and during later stages there are simply less primordial planetesimals left so collisions are just more likely to be major mergers.
	See updated version of figure 9. After including data from more simulations and coloring the points by orbital period, it is clear that there is a difference between the smooth accretion fraction of bodies in the inner disk, compared to the outer disk. In particular, outer disk bodies all have fairly large mass fractions, which implies that they grow in isolation.

The authors state "indicating that accretion of planetesimals is important during all phases of the evolution", yet the smooth accreting fraction is only between 1-10% for all these bodies. Seems to me that for all bodies, regardless of mass and orbital period, major mergers are significantly more important - that is at least with the author's definition of major and minor merger.
	I think we need to loosen our definition what a 'planetesimal' is here. 
	I agree that 1-10% is a small fraction. The mass cutoff should definitely be larger. TODO: Change mass cutoff and hopefully the same scaling with mass remains.
In the legend, I think "orbit" should be put in the subscript instead of P_orbit
746: This is just because in the inner disk there are no primordial planetesimals left to build the more massive bodies, right?
	No, the accretion mode should be qualitatively different, even before the planetesimals run out. TODO: Hopefully including more data in the plot and maybe increasing the mass cutoff for 'smooth' accretion will make this more obvious.
750: use /citep[see][]{Kokubo}
	Fixed
756: I agree that the large bodies must've originated due to mergers with other embryos, but this is not really an explanation for why the feeding zone size doesn't match the theoretical value, no? Instead it's the lack of gravitational scattering, as the authors write above. Maybe the authors can clarify specifically why oligarchic growth models predict this feeding zone size, and also why gravitational scattering is suppressed in the inner disk (due to the lack of primordial planetesimals in the inner disk?).
766: "One should note that [...]." This sentence can be deleted in my opinion.
	Removed
Figure 10:
I suggest indicating the boundary between accretion modes in this figure. >60d orbits only include 4-5 bodies. However, to me it seems like the shape of the accretion zone is qualitatively the same until ~ 20 orbits, which strengthens my view that the 60d orbit boundary is not very clear at all.
I suggest matching the x-axis limits the shown xdata.
I think "Fraction of planetesimals accreted" is a somewhat misleading description of what the plot is showing. What about: "Distribution of accreted planetesimals"?
775: The transition between smooth and unimodal seems to be happening at about 20 days to me.
Figure 11
The y-axis label overlap
I think the finding that the crossing of alpha = 0.1 in the bottom panel well matches the crossing of tilde b = 2 sqrt(3) is interesting and worth highlighting more, maybe with some vertical lines? Is there a meaning behind alpha = 0.1, i.e., is this coming from some analytical consideration? See my major comments
	TODO: Include more sims in top panel of figure 11?

### 2023-04-21 main.tex
Add definition of hill sphere and hill eccentricity

We begin our analysis by considering a disk of equal mass planetesimals
with radius $r_{pl}$, mass $m_{pl}$ and surface density
$\Sigma_{pl}$. The collision rate in the vicinity of an orbit defined
by Keplerian frequency $\Omega$ can be written as $n \Gamma v$, where
$n = \Sigma_{pl} \Omega / 2 m_{pl} v$ (where we have assumed that the scale height of the planetesimal disk goes as $2v/\Omega$). $\Gamma$ describes the effective
collision cross section and $v$ is the typical encounter velocity
between planetesimals.
For a swarm of planetesimals on randomly oriented orbits, $v$ is typically
taken to the rms velocity, which can be related to the eccentricity and inclination distribution $(e, i)$ in the following way \citep{lissauer93}:
\begin{equation}\label{eq:ecc_vel}
	\left< v^{2} \right>^{1/2} = \left( \frac{5}{4} \left< e^{2} \right>^{1/2} + \left< i^{2} \right>^{1/2}  \right) v_{k}.
\end{equation}
Assuming that every collision results in a perfect merger, the growth rate of a planetesimal is given by
\begin{equation}\label{eq:growth}
	\frac{1}{M}\frac{dM}{dt} = \frac{\Sigma \Omega}{2 m_{pl}} \Gamma.
\end{equation}

@@

In figure \ref{fig:timescales}, we show the ratio between the relaxation
and collision timescale for a population of equal-mass planetesimals
as a function of orbital period. Here, the encounter velocity is
described by equation \ref{eq:ecc_vel}. For
simplicity, we assume that $\left< e^2 \right>^{1/2} = 2\left< i^2
\right>^{1/2}$ \citep{ida93a} and that the eccentricity dispersion is
constant with orbital period. The eccentricity dispersion is 
scaled by the Hill factor $\left( m_{pl}/ 3 M_{\star}
\right)^{1/3}$ such that $e_{h} = 1$ corresponds to the boundary
between shear and dispersion dominated encounters. The horizontal
dashed line indicates where $t_{relax} = t_{coll}$. The timescale
criterion for oligarchic growth is only satisfied in regions where the
disk is sufficiently dynamically cold and the orbital period is
sufficiently long. In sections \ref{sec:narrow} and \ref{sec:fulldisk}
we will explore the behavior and outcome of planetesimal accretion in regions where this criterion is \textit{not} satisfied.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/minor_frac_100m.png}
    \caption{For the 100 most massive bodies at the end of the fdHi simulation, the fraction of their total mass attained through mergers with initial mass planetesimals (smooth accretion) as a function of total mass. Bodies that reside beyond the 60 day accretion boundary are shown with red crosses, while bodies interior to the boundary are shown with blue dots. For the short period bodies, there is a decreasing correlation between smooth accretion fraction and mass, which suggests that most growth occurs between equal mass bodies. For the bodies in the outer disk, this trend is flat, indicating that accretion of planetesimals is important during all phases of evolution.\label{fig:minor_frac}}
\end{center}
\end{figure}

>>>>

We begin our analysis by considering a disk of equal mass planetesimals
with radius $r_{pl}$, mass $m_{pl}$ and surface density
$\Sigma_{pl}$. The collision rate in the vicinity of an orbit defined
by Keplerian frequency $\Omega$ can be written as $n \Gamma v$, where
$n = \Sigma_{pl} \Omega / 2 m_{pl} v$ (where we have assumed that the scale height of the planetesimal disk goes as $2v/\Omega$). $\Gamma$ describes the effective
collision cross section and $v$ is the typical encounter velocity
between planetesimals.
For a swarm of planetesimals on randomly oriented orbits, $v$ is typically
taken to the rms velocity, which can be related to the eccentricity and inclination distribution $(e, i)$ in the following way \citep{lissauer93}:
\begin{equation}\label{eq:ecc_vel}
	\left< v^{2} \right>^{1/2} = \left( \frac{5}{4} \left< e^{2} \right>^{1/2} + \left< i^{2} \right>^{1/2}  \right) v_{k}.
\end{equation}

\textbf{The dynamical interactions between growing planetesimals can be somewhat simplified by scaling the orbital elements of the bodies by
a Hill factor}

\begin{equation}\label{eq:hillfac}
	h = \left(\frac{m_{pl}}{3 M_{*}}\right)^{1/3}, 
\end{equation}

\textbf{\noindent where $M_{*}$ is the mass of the central star. With this definition, $r_{h} = h a$ is the Hill radius of a body, which is the distance where the gravity of the growing planetesimal dominates over the gravity of the star. Using equation \ref{eq:hillfac}, the eccentricity, inclination and separation between orbiting bodies can be defined as}

\begin{equation}\label{eq:hillorb}
	e_{h} = \frac{a e}{h}, \: i_{h} = \frac{a i}{h}, \: b = \frac{a_{2} - a_{1}}{h}.
\end{equation}

\textbf{Using this formalism, $e_{h}$ and $i_{h}$ describe the radial and vertical excursions of an orbiting body in units of its Hill radius. For $e_{h} > 1$, the random velocity dispersion dominates over the shearing motion over a separation of $1 r_{h}$ and encounters can be treated with a two-body formalism.}

Assuming that every collision results in a perfect merger, the growth rate of a planetesimal is given by
\begin{equation}\label{eq:growth}
	\frac{1}{M}\frac{dM}{dt} = \frac{\Sigma \Omega}{2 m_{pl}} \Gamma.
\end{equation}

@@

In figure \ref{fig:timescales}, we show the ratio between the relaxation
and collision timescale for a population of equal-mass planetesimals
as a function of orbital period. Here, the encounter velocity is
described by equation \ref{eq:ecc_vel}. For
simplicity, we assume that $\left< e^2 \right>^{1/2} = 2\left< i^2
\right>^{1/2}$ \citep{ida93a} and that the eccentricity dispersion is
constant with orbital period. The horizontal
dashed line indicates where $t_{relax} = t_{coll}$. The timescale
criterion for oligarchic growth is only satisfied in regions where the
disk is sufficiently dynamically cold and the orbital period is
sufficiently long. In sections \ref{sec:narrow} and \ref{sec:fulldisk}
we will explore the behavior and outcome of planetesimal accretion in regions where this criterion is \textit{not} satisfied.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/minor_frac.png}
    \caption{For the 100 most massive bodies at the end of the fdHi simulation, the fraction of their total mass attained through mergers with initial mass planetesimals (smooth accretion) as a function of total mass. Bodies that reside beyond the 60 day accretion boundary are shown with red crosses, while bodies interior to the boundary are shown with blue dots. For the short period bodies, there is a decreasing correlation between smooth accretion fraction and mass, which suggests that most growth occurs between equal mass bodies. For the bodies in the outer disk, this trend is flat, indicating that accretion of planetesimals is important during all phases of evolution.\label{fig:minor_frac}}
\end{center}
\end{figure}
